---
title: "Data_analysis"
author: "Justice Akuoko-Frimpong, Jonathan Ta, Edward Shao"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
library(data.table)
library(here)
load(here("data", "ASTER.RData"))
```

## Exploratory data analysis

```{r, warning=FALSE}
library(viridis)    # for the viridis color scale
library(geoR)       # for calculating the empirical variogram
library(gridExtra)  # for arranging plots

# Spatial distribution of temperature
library(ggplot2)
ggplot(dt, aes(lon, lat)) +
  geom_tile(aes(fill = emis)) + 
  xlab("longitude") +
  ylab("latitude") +
  guides(fill = guide_legend(title = "emissivity")) +
  ggtitle("Spatial emissivity Distribution") +
  coord_fixed() 

ggplot(dt, aes(lon, lat)) +
  geom_tile(aes(fill = ndvi)) + 
  xlab("longitude") +
  ylab("latitude") +
  guides(fill = guide_legend(title = "NDVI")) +
  ggtitle("Spatial ndvi Distribution") +
  coord_fixed() 

ggplot(dt, aes(lon, lat)) +
  geom_tile(aes(fill = temp)) + 
  xlab("longitude") +
  ylab("latitude") +
  guides(fill = guide_legend(title = "temperature")) +
  ggtitle("Spatial temperature Distribution") +
  coord_fixed()

```
## Semivariograms
```{r}
# subset data further for more efficient semivariogram analysis
lat.sub = dt[, unique(lat)]
lat.sub = lat.sub[seq(1, length(lat.sub), by = 2)]

lon.sub = dt[, unique(lon)]
lon.sub = lon.sub[seq(1, length(lon.sub), by = 2)]

dt.sub = dt[lat %in% lat.sub & lon %in% lon.sub]
dt.sub = dt.sub[complete.cases(dt.sub[, c("lat", "lon", "ndvi", "temp")]), ]
```


```{r}
# semivariogram of emissivity
sv = variog(
  coords = cbind(dt.sub[!is.na(emis), lat], dt.sub[!is.na(emis), lon]),
  data = dt.sub[!is.na(emis), emis]
  )

sv.dt = data.table(
  dists = sv$u,
  variogram = sv$v,
  npairs = sv$n,
  sd = sv$sd
)

ggplot(sv.dt, aes(x = dists, y = variogram)) + 
  geom_point() + 
  labs(title = "emissivity semivariogram")



# identifying min and max distances
summary(sv.dt)
```

```{r}
# semivariogram of NDVI
sv = variog(
  coords = cbind(dt.sub[, lat], dt.sub[, lon]),
  data = dt.sub[, ndvi]
  )

sv.dt = data.table(
  dists = sv$u,
  variogram = sv$v,
  npairs = sv$n,
  sd = sv$sd
)
ggplot(sv.dt, aes(x = dists, y = variogram)) + 
  geom_point() + 
  labs(title = "NDVI semivariogram")


# identifying min and max distances
summary(sv.dt)
```

```{r}
# semivariogram of temperature
sv = variog(
  coords = cbind(dt.sub[, lat], dt.sub[, lon]),
  data = dt.sub[, temp]
  )

sv.dt = data.table(
  dists = sv$u,
  variogram = sv$v,
  npairs = sv$n,
  sd = sv$sd
)
 ggplot(sv.dt, aes(x = dists, y = variogram)) + 
  geom_point() + 
  labs(title = "temperature semivariogram")

# identifying min and max distances
summary(sv.dt)
```


## Data preparation
```{r}
load(here("data", "ASTER.RData"))
```

```{r}
library(svc)

complete_cases <- complete.cases(dt)
dt_complete <- dt[complete_cases, ]

# Create standardized inputs
Y <- dt_complete$temp
X <- scale(as.matrix(dt_complete[, c("ndvi", "emis")]))  
s <- scale(as.matrix(dt_complete[, c("lon", "lat")]))    

# Knot creation function with checks
create_knots <- function(coords, k = 50) {
  if(!all(c("lat", "lon") %in% colnames(coords))) {
    stop("coords must contain 'lat' and 'lon' columns")
  }
  
  lat_knots <- unique(na.omit(coords[,"lat"]))
  lon_knots <- unique(na.omit(coords[,"lon"]))
  
  lat_knots <- lat_knots[seq(1, length(lat_knots), by = k)]
  lon_knots <- lon_knots[seq(1, length(lon_knots), by = k)]
  
  knots <- as.matrix(expand.grid(lat_knots, lon_knots))
  colnames(knots) <- c("lat", "lon")
  return(knots)
}

# Create and validate knots
coords <- as.data.frame(dt_complete[, c("lat", "lon")])
knots <- create_knots(coords, k = 50)
knots_df <- merge(data.frame(knots), dt_complete, by = c("lat", "lon"), all.x = TRUE)

complete_knots <- complete.cases(knots_df[, c("temp", "ndvi", "emis")])

# Filter all components to only keep complete cases
knots_df <- knots_df[complete_knots, ]
Y_knots <- knots_df$temp
X_knots <- as.matrix(knots_df[, c("ndvi", "emis")])
knots_matrix <- as.matrix(knots_df[, c("lon", "lat")])

# Now scale the complete data
X_knots <- scale(X_knots)
knots_matrix <- scale(knots_matrix)
```

### Model initilizations
```{r}
# Initial values with small phi's for stability
init_values <- list(
  beta_knots_start = matrix(rnorm(nrow(knots_matrix)*ncol(X), sd = 0.1), 
                          nrow = nrow(knots_matrix), ncol = ncol(X)),
  w_knots_start = rnorm(nrow(knots_matrix), sd = 0.1),
  phi_beta_start = rep(0.01, ncol(X)),  # Very small initial phi values
  phi_w_start = 0.01,
  sigmasq_beta_start = rep(0.1, ncol(X)),
  sigmasq_w_start = 0.1,
  tausq_start = 0.001
)

# bounds and small proposal sizes for phi's
tuning_params <- list(
  phi_beta_proposal_sd = rep(0.005, ncol(X)),  # Small steps
  phi_w_proposal_sd = 0.005,
  lower_beta = rep(0.001, ncol(X)),  # Tight bounds
  upper_beta = rep(0.5, ncol(X)),    
  lower_w = 0.001,
  upper_w = 0.5
)

# Weakly informative priors
priors <- list(
  a_beta = rep(2, ncol(X)),  # Slightly informative
  b_beta = rep(1, ncol(X)),
  a_w = 2,
  b_w = 1,
  a_t = 2,
  b_t = 1
)
```


### Model running
```{r, echo=FALSE }
# Run model with error handling
results <- GP_Gibbs(
    Y = Y,
    X = X,
    s = s,
    knots = knots_matrix,
    Y_knots = Y_knots,
    X_knots = X_knots,
    beta_knots_start = init_values$beta_knots_start,
    w_knots_start = init_values$w_knots_start,
    phi_beta_start = init_values$phi_beta_start,
    phi_w_start = init_values$phi_w_start,
    sigmasq_beta_start = init_values$sigmasq_beta_start,
    sigmasq_w_start = init_values$sigmasq_w_start,
    tausq_start = init_values$tausq_start,
    phi_beta_proposal_sd = tuning_params$phi_beta_proposal_sd,
    phi_w_proposal_sd = tuning_params$phi_w_proposal_sd,
    a_beta = priors$a_beta,
    b_beta = priors$b_beta,
    a_w = priors$a_w,
    b_w = priors$b_w,
    a_t = priors$a_t,
    b_t = priors$b_t,
    lower_beta = tuning_params$lower_beta,
    upper_beta = tuning_params$upper_beta,
    lower_w = tuning_params$lower_w,
    upper_w = tuning_params$upper_w,
    mcmc = 1000  
  )


```


###  Trace plots

```{r}
library(ggplot2)
library(tidyr)

# Convert samples to data frame
trace_data <- data.frame(
  iteration = 1:length(results$phi_w_samples),
  phi_beta_ndvi = results$phi_beta_samples[,1],
  phi_beta_emis = results$phi_beta_samples[,2],
  phi_w = results$phi_w_samples,
  sigmasq_beta_ndvi = results$sigmasq_beta_samples[,1],
  sigmasq_beta_emis = results$sigmasq_beta_samples[,2],
  sigmasq_w = results$sigmasq_w_samples,
  tausq = results$tausq_samples
)

# Create faceted trace plots
trace_long <- trace_data %>%
  pivot_longer(-iteration, names_to = "parameter", values_to = "value")

ggplot(trace_long, aes(x = iteration, y = value)) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ parameter, scales = "free_y", ncol = 2) +
  labs(title = "MCMC Trace Plots", x = "Iteration", y = "Parameter Value") +
  theme_minimal()

```
### plots of beta at selected locations

```{r}
# Trace plots for NDVI coefficient at selected locations
plot(results$beta_samples[5, 1, ], type = 'l', 
     main = "Beta NDVI - Location 1", ylab = "Coefficient Value")
plot(results$beta_samples[10, 1, ], type = 'l', 
     main = "Beta NDVI - Location 10", ylab = "Coefficient Value")

# Trace plots for Emissivity coefficient at selected locations
plot(results$beta_samples[1, 2, ], type = 'l', 
     main = "Beta Emissivity - Location 1", ylab = "Coefficient Value")
plot(results$beta_samples[10, 2, ], type = 'l', 
     main = "Beta Emissivity - Location 10", ylab = "Coefficient Value")

```

### Interpreations and Problems

> 1.) The covariate effects converged near zero, suggesting weak linear relationships between NDVI/emissivity and temperature in this system
2.) The large estimated variance components  indicate substantial unexplained spatial autocorrelation


Together, these results suggest:

1. Identifiability Problem:
-  The spatial random effects (w) may be absorbing all the spatial signal
- This leaves little variation for the covariate effects to explain

2. Problem with some specifications

Next steps:

1. Recode our function to perform similarly to the spBayes package's spatially varying coefficient (SVC) model