---
title: "Untitled"
author: "new"
date: "4/4/2025"
output: pdf_document
---

### Initialization
```{r}
library(knitr)
library(varycoef)
library(spBayes)
library(ggplot2)
library(gridExtra)
library(Matrix)
library(here)
set.seed(123)

calc_C_phi = function(coords, phi) {
  n = nrow(coords)
  dists = as.matrix(dist(coords)) ^ 2
  C_phi = exp(-0.5 / phi * dists)
}

mcmc = 1000

```

### Purpose
# The purpose of this simulation study is compare the performance between our method
# and varycoef R package. Specfically, we will assess the bias, computational speed, average length of credible interval
# and coverage

## Setting
# 5 replications (due to time constraint)
# We will perform 1000 iterations of mcmc 
# We will vary by 100 or 500 sample size
# We will vary by number of betas (2 or 3) and priors (low or high)


## metric
### Bias ###
### 1/R (\sum_{r=1} (beta hat - true beta ))  


# How long to run in seconds

#### Simulations
```{r}

set.seed(123)
svclm_full = data.frame()
svclm_low = data.frame()
coef = data.frame()
bayes = data.frame()
for(i in 1:1000)
{
  lat = seq(0, 20, by = 1)
  lon = seq(0, 20, by = 1)
  
  coords = as.matrix(expand.grid(lat, lon))
  colnames(coords) = c("lat", "lon")
  
  n = nrow(coords)
  p = 2
  
  # generating w
  sigmasq_w = 0.1
  phi_w = 2
  
  C_w = calc_C_phi(coords, phi_w)
  w = MASS::mvrnorm(1, rep(0, n), sigmasq_w * C_w)
  
  sigmasq_1 = 0.1
  phi_1 = 2
  
  C_1 = calc_C_phi(coords, phi_1)
  beta_1_mean = 3
  beta_1 = MASS::mvrnorm(1, rep(beta_1_mean, n),sigmasq_1 * C_1 )
  
  sigmasq_2 = 0.1
  phi_2 = 2
  
  C_2 = calc_C_phi(coords, phi_2)
  beta_2_mean = -5
  beta_2 = MASS::mvrnorm(1, rep(beta_2_mean, n),sigmasq_2 * C_2)
  
  # generating X
  X_1 = rnorm(n, 0, 1)
  X_2 = rnorm(n, 5, 1)
  
  # generating epsilon
  tausq = 0.0001
  epsilon = rnorm(n, mean = 0, sd = sqrt(tausq))
  
  Y = X_1 * beta_1 + X_2 * beta_2 + w + epsilon
  
  
  X = cbind(X_1,X_2)
  X = cbind(1,X)
  m = nrow(knots)
  l = 0
  u = 4
  
  start_time <- Sys.time()
  full_rank = svclm(Y=Y, X=X,coords= coords,phi_lower = rep(l,3),
                         phi_upper =rep(u,3))
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  
  svclm_full = rbind(svclm_full,c(mean(colMeans(full_rank$w_samples[800:1000, ,1])),mean(colMeans(full_rank$w_samples[800:1000, ,2])),mean(colMeans(full_rank$w_samples[800:1000, ,3])), time_taken ))
  
 knot_results = simpleknots(
  Y = Y,
  X = X,
  coords = coords,
  k = 2  
)

# Extract knot components
Y_star = knot_results$Y_knots
X_star = knot_results$X_knots
knots = knot_results$knots

  
  m = nrow(knots)
  
  start_time <- Sys.time()
  low_rank = svclm(Y=Y, X=X, coords = coords,Y_knots = Y_star,
                         X_knots = X_star,
                         knots = knots,phi_lower = rep(l,3),
                         phi_upper =rep(u,3))
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  
   svclm_low = rbind(svclm_low,c(mean(colMeans(low_rank$w_samples[800:1000, ,1])),mean(colMeans(low_rank$w_samples[800:1000, ,2])),mean(colMeans(low_rank$w_samples[800:1000, ,3])), time_taken ))
  
  
  
  
  control <- SVC_mle_control(
  # initial values of optimization
  init = rep(0.1, 2*3+1),
  # using profile likelihood
  profileLik = TRUE
)

# controls specific to optimization procedure, see help(optim)
opt.control <- list(
  # number of iterations (set to one for demonstration sake)
  maxit = 1,
  # tracing information
  trace = 6
)
  start_time <- Sys.time()
  fit <- SVC_mle(y = Y, X = X,locs = coords,control = control,
               optim.control = opt.control)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
 coef = rbind(coef,c(fit$coefficients[1],fit$coefficients[2],fit$coefficients[3],time_taken)) 
  
  start_time <- Sys.time()
  svc_fit <- spSVC(
  y ~ x1 + x2,
  coords = coords,
  data = data.frame(y = Y, x1 = X_1, x2 = X_2),
  starting = list(
    "phi" = rep(3, 3),
    "sigma.sq" = rep(1, 3),
    "tau.sq" = 0.01
  ),
  tuning = list(
    "phi" = rep(0.1, 3),
    "sigma.sq" = rep(0.1, 3),
    "tau.sq" = 0.01
  ),
  priors = list(
    "phi.Unif" = list(rep(0.1, 3), rep(10, 3)),
    "sigma.sq.IG" = list(rep(2, 3), rep(2, 3)),
    "tau.sq.IG" = list(2, 0.01)
  ),
  cov.model = "exponential",
  n.samples = 1000, svc.cols= c(1,2,3)
)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  
  svc_fit_rec <- spRecover(svc_fit, start = 800, verbose = FALSE,get.beta= TRUE,get.w= TRUE)
  beta_samples <- svc_fit_rec$p.beta.recover.samples
  w_posterior_samples <- as.matrix(svc_fit_rec$p.w.recover.samples)
  w_means <- mean(colMeans(w_posterior_samples))
  bayes = rbind(bayes,c(w_means,mean(beta_samples[,1]),mean(beta_samples[,2]),time_taken))
}
```































```{r,warning=FALSE}
### Scenario 1: (2 betas and "low" priors)
### 100 Sample size
gibbs1 = data.frame()
coef1 = data.frame()
for (i in 1:1000) {
  
  lat = seq(0, 9, by = 1)
  lon = seq(0, 9, by = 1)
  
  coords = as.matrix(expand.grid(lat, lon))
  colnames(coords) = c("lat", "lon")
  
  n = nrow(coords)
  p = 2
  
  # generating w
  sigmasq_w = 1
  phi_w = 2
  
  C_w = calc_C_phi(coords, phi_w)
  w = MASS::mvrnorm(1, rep(0, n), sigmasq_w * C_w)
  
  # generating beta's
  sigmasq_1 = 1
  phi_1 = 2
  
  C_1 = calc_C_phi(coords, phi_1)
  beta_1_mean = 3
  beta_1 = MASS::mvrnorm(1, rep(beta_1_mean, n),sigmasq_1 * C_1 )
  
  sigmasq_2 = 1
  phi_2 = 2
  
  C_2 = calc_C_phi(coords, phi_2)
  beta_2_mean = -5
  beta_2 = MASS::mvrnorm(1, rep(beta_2_mean, n),sigmasq_2 * C_2)
  
  # generating X
  X_1 = rnorm(n, 0, 10)
  X_2 = rnorm(n, 10, 100)
  
  # generating epsilon
  tausq = 0.0001
  epsilon = rnorm(n, mean = 0, sd = sqrt(tausq))
  
  Y = X_1 * beta_1 + X_2 * beta_2 + w + epsilon
  
  k = 2
  
  lat_knots = unique(lat)
  lat_knots = lat_knots[seq(1, length(lat_knots), by = k)]
  
  lon_knots = unique(lon)
  lon_knots = lon_knots[seq(1, length(lon_knots), by = k)]
  
  knots = as.matrix(expand.grid(lat_knots, lon_knots))
  
  df = data.frame(coords, Y, X_1, X_2, w, beta_1, beta_2, epsilon)
  
  knots_df = data.frame(knots)
  colnames(knots_df) = c("lat", "lon")
  
  knots_df = merge(knots_df, df, by = c("lat", "lon"), all.x = TRUE, all.y = FALSE)
  
  X = as.matrix(df[, c("X_1", "X_2")])
  Y_star = knots_df$Y
  X_star = as.matrix(knots_df[, c("X_1", "X_2")])
  
  m = nrow(knots)
  l = 0
  u = 4
  start_time <- Sys.time()
    output = svc::svclm_beta(Y = Y,
                         X = X,
                         s = coords,
                         Y_knots = Y_star,
                         X_knots = X_star,
                         knots = knots,
                         beta_knots_start = matrix(0, nrow = m, ncol = p),
                         phi_beta_start = rep(l + (u - l) / 2, p),
                         sigmasq_beta_start = rep(1, p),
                         tausq_start = 1,
                         phi_beta_proposal_sd = rep(0.5, p),
                         phi_beta_lower = rep(l, p),
                         phi_beta_upper = rep(u, p),
                         mcmc = mcmc)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  gibbs1 = rbind(gibbs1,c(mean(colMeans(output$beta_samples[800:1000, ,1]) -3),mean(colMeans(output$beta_samples[800:1000, ,2]) +5), time_taken ))
  X = cbind(1,X)
  start_time <- Sys.time()
  fit <- SVC_mle(y = Y, X = X,locs = coords)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  
  coef1 = rbind(coef1,c(fit$coefficients[2]-3,fit$coefficients[3]+5,time_taken))
  }


gibbs2 = data.frame()
coef2 = data.frame()
### 361 Sample size
for (i in 1:5) {
  lat = seq(0, 9, by = .4)
  lon = seq(0, 9, by = .4)
  
  coords = as.matrix(expand.grid(lat, lon))
  colnames(coords) = c("lat", "lon")
  
  n = nrow(coords)
  p = 2
  
  # generating w
  sigmasq_w = 1
  phi_w = 2
  C_w = calc_C_phi(coords, phi_w)
  w = MASS::mvrnorm(1, rep(0, n), sigmasq_w * C_w)
  
  # generating beta's
  sigmasq_1 = 1
  phi_1 = 2
  C_1 = calc_C_phi(coords, phi_1)
  beta_1_mean = 3
  beta_1 = MASS::mvrnorm(1, rep(beta_1_mean, n), sigmasq_1 * C_1)
  
  sigmasq_2 = 1
  phi_2 = 2
  C_2 = calc_C_phi(coords, phi_2)
  beta_2_mean = -5
  beta_2 = MASS::mvrnorm(1, rep(beta_2_mean, n), sigmasq_2 * C_2)
  
  # generating X
  X_1 = rnorm(n, 0, 10)
  X_2 = rnorm(n, 10, 100)
  
  # generating epsilon
  tausq = 0.0001
  epsilon = rnorm(n, mean = 0, sd = sqrt(tausq))
  
  Y = X_1 * beta_1 + X_2 * beta_2 + w + epsilon
  
  k = 2
  
  lat_knots = unique(lat)
  lat_knots = lat_knots[seq(1, length(lat_knots), by = k)]
  
  lon_knots = unique(lon)
  lon_knots = lon_knots[seq(1, length(lon_knots), by = k)]
  
  knots = as.matrix(expand.grid(lat_knots, lon_knots))
  
  df = data.frame(coords, Y, X_1, X_2, w, beta_1, beta_2, epsilon)
  
  knots_df = data.frame(knots)
  colnames(knots_df) = c("lat", "lon")
  
  knots_df = merge(knots_df, df, by = c("lat", "lon"), all.x = TRUE, all.y = FALSE)
  
  X = as.matrix(df[, c("X_1", "X_2")])
  Y_star = knots_df$Y
  X_star = as.matrix(knots_df[, c("X_1", "X_2")])
  
  m = nrow(knots)
  l = 0
  u = 4
  
  start_time <- Sys.time()
output = svc::svclm_beta(Y = Y,
                         X = X,
                         s = coords,
                         Y_knots = Y_star,
                         X_knots = X_star,
                         knots = knots,
                         beta_knots_start = matrix(0, nrow = m, ncol = p),
                         phi_beta_start = rep(l + (u - l) / 2, p),
                         sigmasq_beta_start = rep(1, p),
                         tausq_start = 1,
                         phi_beta_proposal_sd = rep(0.5, p),
                         phi_beta_lower = rep(l, p),
                         phi_beta_upper = rep(u, p),
                         mcmc = mcmc)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  gibbs2 = rbind(gibbs2,c(mean(colMeans(output$beta_samples[800:1000, ,1]) -3),mean(colMeans(output$beta_samples[800:1000, ,2]) +5), time_taken ))
  X = cbind(1,X)
  start_time <- Sys.time()
 fit <- SVC_mle(y = Y, X = X,locs = coords)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  
  coef2 = rbind(coef2,c(fit$coefficients[2]-3,fit$coefficients[3]+5,time_taken))
}
```






```{r,warning=FALSE}
gibbs3 = data.frame()
coef3 = data.frame()
### Scenario 2: (2 betas and "high" priors)
### 100 Sample size
for (i in 1:5) {
  
  lat = seq(0, 9, by = 1)
  lon = seq(0, 9, by = 1)
  
  coords = as.matrix(expand.grid(lat, lon))
  colnames(coords) = c("lat", "lon")
  
  n = nrow(coords)
  p = 2
  
  # generating w
  sigmasq_w = 4
  phi_w = 8
   
  C_w = calc_C_phi(coords, phi_w)
  w = MASS::mvrnorm(1, rep(0, n), sigmasq_w * C_w)
  
  # generating beta's
  sigmasq_1 = 4
  phi_1 = 8
   
  C_1 = calc_C_phi(coords, phi_1)
  beta_1_mean = 3
  beta_1 = MASS::mvrnorm(1, rep(beta_1_mean, n), sigmasq_1 * C_1)
  
  sigmasq_2 = 4
  phi_2 = 8
   
  C_2 = calc_C_phi(coords, phi_2)
  beta_2_mean = -5
  beta_2 = MASS::mvrnorm(1, rep(beta_2_mean, n), sigmasq_2 * C_2)
  
  # generating X
  X_1 = rnorm(n, 0, 10)
  X_2 = rnorm(n, 10, 100)
  
  # generating epsilon
  tausq = 0.01
  epsilon = rnorm(n, mean = 0, sd = sqrt(tausq))
  
  Y = X_1 * beta_1 + X_2 * beta_2 + w + epsilon
  
  k = 2
  
  lat_knots = unique(lat)
  lat_knots = lat_knots[seq(1, length(lat_knots), by = k)]
  
  lon_knots = unique(lon)
  lon_knots = lon_knots[seq(1, length(lon_knots), by = k)]
  
  knots = as.matrix(expand.grid(lat_knots, lon_knots))
  
  df = data.frame(coords, Y, X_1, X_2, w, beta_1, beta_2, epsilon)
  
  knots_df = data.frame(knots)
  colnames(knots_df) = c("lat", "lon")
  
  knots_df = merge(knots_df, df, by = c("lat", "lon"), all.x = TRUE, all.y = FALSE)
  
  X = as.matrix(df[, c("X_1", "X_2")])
  Y_star = knots_df$Y
  X_star = as.matrix(knots_df[, c("X_1", "X_2")])
  
  m = nrow(knots)
  l = 0
  u = 4
  
  start_time <- Sys.time()
 output = svc::svclm_beta(Y = Y,
                         X = X,
                         s = coords,
                         Y_knots = Y_star,
                         X_knots = X_star,
                         knots = knots,
                         beta_knots_start = matrix(0, nrow = m, ncol = p),
                         phi_beta_start = rep(l + (u - l) / 2, p),
                         sigmasq_beta_start = rep(1, p),
                         tausq_start = 1,
                         phi_beta_proposal_sd = rep(0.5, p),
                         phi_beta_lower = rep(l, p),
                         phi_beta_upper = rep(u, p),
                         mcmc = mcmc)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  gibbs3 = rbind(gibbs3,c(mean(colMeans(output$beta_samples[800:1000, ,1]) -3),mean(colMeans(output$beta_samples[800:1000, ,2]) +5), time_taken ))
  X = cbind(1,X)
  start_time <- Sys.time()
 fit <- SVC_mle(y = Y, X = X,locs = coords)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  
  coef3 = rbind(coef3,c(fit$coefficients[2]-3,fit$coefficients[3]+5,time_taken))
}

gibbs4 = data.frame()
coef4 = data.frame()
### 361 Sample size
for (i in 1:5) {
  lat = seq(0, 9, by = .4)
  lon = seq(0, 9, by = .4)
  
  coords = as.matrix(expand.grid(lat, lon))
  colnames(coords) = c("lat", "lon")
  
  n = nrow(coords)
  p = 2
  
  # generating w
  sigmasq_w = 4
  phi_w = 8
  C_w = calc_C_phi(coords, phi_w)
  w = MASS::mvrnorm(1, rep(0, n), sigmasq_w * C_w)
  
  # generating beta's
  sigmasq_1 = 4
  phi_1 = 8
  C_1 = calc_C_phi(coords, phi_1)
  beta_1_mean = 3
  beta_1 = MASS::mvrnorm(1, rep(beta_1_mean, n), sigmasq_1 * C_1)
  
  sigmasq_2 = 4
  phi_2 = 8
  C_2 = calc_C_phi(coords, phi_2)
  beta_2_mean = -5
  beta_2 = MASS::mvrnorm(1, rep(beta_2_mean, n), sigmasq_2 * C_2)
  
  # generating X
  X_1 = rnorm(n, 0, 10)
  X_2 = rnorm(n, 10, 100)
  
  # generating epsilon
  tausq = 0.01
  epsilon = rnorm(n, mean = 0, sd = sqrt(tausq))
  
  Y = X_1 * beta_1 + X_2 * beta_2 + w + epsilon
  
  k = 2
  
  lat_knots = unique(lat)
  lat_knots = lat_knots[seq(1, length(lat_knots), by = k)]
  
  lon_knots = unique(lon)
  lon_knots = lon_knots[seq(1, length(lon_knots), by = k)]
  
  knots = as.matrix(expand.grid(lat_knots, lon_knots))
  
  df = data.frame(coords, Y, X_1, X_2, w, beta_1, beta_2, epsilon)
  
  knots_df = data.frame(knots)
  colnames(knots_df) = c("lat", "lon")
  
  knots_df = merge(knots_df, df, by = c("lat", "lon"), all.x = TRUE, all.y = FALSE)
  
  X = as.matrix(df[, c("X_1", "X_2")])
  Y_star = knots_df$Y
  X_star = as.matrix(knots_df[, c("X_1", "X_2")])
  
  m = nrow(knots)
  l = 0
  u = 4
  
  start_time <- Sys.time()
output = svc::svclm_beta(Y = Y,
                         X = X,
                         s = coords,
                         Y_knots = Y_star,
                         X_knots = X_star,
                         knots = knots,
                         beta_knots_start = matrix(0, nrow = m, ncol = p),
                         phi_beta_start = rep(l + (u - l) / 2, p),
                         sigmasq_beta_start = rep(1, p),
                         tausq_start = 1,
                         phi_beta_proposal_sd = rep(0.5, p),
                         phi_beta_lower = rep(l, p),
                         phi_beta_upper = rep(u, p),
                         mcmc = mcmc)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  gibbs4 = rbind(gibbs4,c(mean(colMeans(output$beta_samples[800:1000, ,1]) -3),mean(colMeans(output$beta_samples[800:1000, ,2]) +5), time_taken ))
  X = cbind(1,X)
  start_time <- Sys.time()
 fit <- SVC_mle(y = Y, X = X,locs = coords)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  
  coef4 = rbind(coef4,c(fit$coefficients[2]-3,fit$coefficients[3]+5,time_taken))
}

```





```{r,warning=FALSE}
gibbs5 = data.frame()
coef5 = data.frame()
### Scenario 3: (3 betas and "low" priors)
### 100 Sample size
for (i in 1:5) {
  
  lat = seq(0, 9, by = 1)
  lon = seq(0, 9, by = 1)
  
  coords = as.matrix(expand.grid(lat, lon))
  colnames(coords) = c("lat", "lon")
  
  n = nrow(coords)
  p = 3
  
  # generating w
  sigmasq_w = 1
  phi_w = 2
   
  C_w = calc_C_phi(coords, phi_w)
  w = MASS::mvrnorm(1, rep(0, n), sigmasq_w * C_w)
  
  # generating beta's
  sigmasq_1 = 1
  phi_1 = 2
   
  C_1 = calc_C_phi(coords, phi_1)
  beta_1_mean = 3
  beta_1 = MASS::mvrnorm(1, rep(beta_1_mean, n), sigmasq_1 * C_1)
  
  sigmasq_2 = 1
  phi_2 = 2
   
  C_2 = calc_C_phi(coords, phi_2)
  beta_2_mean = -5
  beta_2 = MASS::mvrnorm(1, rep(beta_2_mean, n), sigmasq_2 * C_2)
  
  sigmasq_3 = 1
  phi_3 = 2
   
  C_3 = calc_C_phi(coords, phi_3)
  beta_3_mean = -10
  beta_3 = MASS::mvrnorm(1, rep(beta_3_mean, n), sigmasq_3 * C_3)
  
  # generating X
  X_1 = rnorm(n, 0, 10)
  X_2 = rnorm(n, 10, 100)
  X_3 = rnorm(n, 100, 1000)
  
  # generating epsilon
  tausq = 0.0001
  epsilon = rnorm(n, mean = 0, sd = sqrt(tausq))
  
  Y = X_1 * beta_1 + X_2 * beta_2+ X_3 * beta_3 + w + epsilon
  
  k = 2
  
  lat_knots = unique(lat)
  lat_knots = lat_knots[seq(1, length(lat_knots), by = k)]
  
  lon_knots = unique(lon)
  lon_knots = lon_knots[seq(1, length(lon_knots), by = k)]
  
  knots = as.matrix(expand.grid(lat_knots, lon_knots))
  
  df = data.frame(coords, Y, X_1, X_2,X_3, w, beta_1, beta_2, beta_3, epsilon)
  
  knots_df = data.frame(knots)
  colnames(knots_df) = c("lat", "lon")
  
  knots_df = merge(knots_df, df, by = c("lat", "lon"), all.x = TRUE, all.y = FALSE)
  
  X = as.matrix(df[, c("X_1", "X_2","X_3")])
  Y_star = knots_df$Y
  X_star = as.matrix(knots_df[, c("X_1", "X_2","X_3")])
  
  m = nrow(knots)
  l = 0
  u = 4
  
  start_time <- Sys.time()
output = svc::svclm_beta(Y = Y,
                         X = X,
                         s = coords,
                         Y_knots = Y_star,
                         X_knots = X_star,
                         knots = knots,
                         beta_knots_start = matrix(0, nrow = m, ncol = p),
                         phi_beta_start = rep(l + (u - l) / 2, p),
                         sigmasq_beta_start = rep(1, p),
                         tausq_start = 1,
                         phi_beta_proposal_sd = rep(0.5, p),
                         phi_beta_lower = rep(l, p),
                         phi_beta_upper = rep(u, p),
                         mcmc = mcmc)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  gibbs5 = rbind(gibbs5,c(mean(colMeans(output$beta_samples[800:1000, ,1]) -3),mean(colMeans(output$beta_samples[800:1000, ,2]) +5),mean(colMeans(output$beta_samples[800:1000, ,3]) +10), time_taken ))
  X = cbind(1,X)
  start_time <- Sys.time()
 fit <- SVC_mle(y = Y, X = X,locs = coords)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  
  coef5 = rbind(coef5,c(fit$coefficients[2]-3,fit$coefficients[3]+5,fit$coefficients[4]+10,time_taken))
}

gibbs6 = data.frame()
coef6 = data.frame()
### 361 Sample size
for (i in 1:5) {
  lat = seq(0, 9, by = .4)
  lon = seq(0, 9, by = .4)
  
  coords = as.matrix(expand.grid(lat, lon))
  colnames(coords) = c("lat", "lon")
  
  n = nrow(coords)
  p = 3
  
  # generating w
  sigmasq_w = 1
  phi_w = 2
   
  C_w = calc_C_phi(coords, phi_w)
  w = MASS::mvrnorm(1, rep(0, n), sigmasq_w * C_w)
  
  # generating beta's
  sigmasq_1 = 1
  phi_1 = 2
   
  C_1 = calc_C_phi(coords, phi_1)
  beta_1_mean = 3
  beta_1 = MASS::mvrnorm(1, rep(beta_1_mean, n), sigmasq_1 * C_1)
  
  sigmasq_2 = 1
  phi_2 = 2
   
  C_2 = calc_C_phi(coords, phi_2)
  beta_2_mean = -5
  beta_2 = MASS::mvrnorm(1, rep(beta_2_mean, n), sigmasq_2 * C_2)
  
  sigmasq_3 = 1
  phi_3 = 2
   
  C_3 = calc_C_phi(coords, phi_3)
  beta_3_mean = -10
  beta_3 = MASS::mvrnorm(1, rep(beta_3_mean, n), sigmasq_3 * C_3)
  
  # generating X
  X_1 = rnorm(n, 0, 10)
  X_2 = rnorm(n, 10, 100)
  X_3 = rnorm(n, 100, 1000)
  
  # generating epsilon
  tausq = 0.0001
  epsilon = rnorm(n, mean = 0, sd = sqrt(tausq))
  
  Y = X_1 * beta_1 + X_2 * beta_2+ X_3 * beta_3 + w + epsilon
  
  k = 2
  
  lat_knots = unique(lat)
  lat_knots = lat_knots[seq(1, length(lat_knots), by = k)]
  
  lon_knots = unique(lon)
  lon_knots = lon_knots[seq(1, length(lon_knots), by = k)]
  
  knots = as.matrix(expand.grid(lat_knots, lon_knots))
  
  df = data.frame(coords, Y, X_1, X_2,X_3, w, beta_1, beta_2, beta_3, epsilon)
  
  knots_df = data.frame(knots)
  colnames(knots_df) = c("lat", "lon")
  
  knots_df = merge(knots_df, df, by = c("lat", "lon"), all.x = TRUE, all.y = FALSE)
  
  X = as.matrix(df[, c("X_1", "X_2","X_3")])
  Y_star = knots_df$Y
  X_star = as.matrix(knots_df[, c("X_1", "X_2","X_3")])
  
  m = nrow(knots)
  l = 0
  u = 4
  
  start_time <- Sys.time()
output = svc::svclm_beta(Y = Y,
                         X = X,
                         s = coords,
                         Y_knots = Y_star,
                         X_knots = X_star,
                         knots = knots,
                         beta_knots_start = matrix(0, nrow = m, ncol = p),
                         phi_beta_start = rep(l + (u - l) / 2, p),
                         sigmasq_beta_start = rep(1, p),
                         tausq_start = 1,
                         phi_beta_proposal_sd = rep(0.5, p),
                         phi_beta_lower = rep(l, p),
                         phi_beta_upper = rep(u, p),
                         mcmc = mcmc)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  gibbs6 = rbind(gibbs6,c(mean(colMeans(output$beta_samples[800:1000, ,1]) -3),mean(colMeans(output$beta_samples[800:1000, ,2]) +5),mean(colMeans(output$beta_samples[800:1000, ,3]) +10), time_taken ))
  X = cbind(1,X)
  start_time <- Sys.time()
 fit <- SVC_mle(y = Y, X = X,locs = coords)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  
  coef6 = rbind(coef6,c(fit$coefficients[2]-3,fit$coefficients[3]+5,fit$coefficients[4]+10,time_taken))
}
```



```{r,warning=FALSE}
gibbs7 = data.frame()
coef7 = data.frame()
### Scenario 4: (3 betas and "high" priors)
### 100 Sample size
for (i in 1:5) {
  
  lat = seq(0, 9, by = 1)
  lon = seq(0, 9, by = 1)
  
  coords = as.matrix(expand.grid(lat, lon))
  colnames(coords) = c("lat", "lon")
  
  n = nrow(coords)
  p = 3
  
  # generating w
  sigmasq_w = 4
  phi_w = 8
   
  C_w = calc_C_phi(coords, phi_w)
  w = MASS::mvrnorm(1, rep(0, n), sigmasq_w * C_w)
  
  # generating beta's
  sigmasq_1 = 4
  phi_1 = 8
   
  C_1 = calc_C_phi(coords, phi_1)
  beta_1_mean = 3
  beta_1 = MASS::mvrnorm(1, rep(beta_1_mean, n), sigmasq_1 * C_1)
  
  sigmasq_2 = 4
  phi_2 = 8
   
  C_2 = calc_C_phi(coords, phi_2)
  beta_2_mean = -5
  beta_2 = MASS::mvrnorm(1, rep(beta_2_mean, n), sigmasq_2 * C_2)
  
  sigmasq_3 = 4
  phi_3 = 8
   
  C_3 = calc_C_phi(coords, phi_3)
  beta_3_mean = -10
  beta_3 = MASS::mvrnorm(1, rep(beta_3_mean, n), sigmasq_3 * C_3)
  
  # generating X
  X_1 = rnorm(n, 0, 10)
  X_2 = rnorm(n, 10, 100)
  X_3 = rnorm(n, 100, 1000)
  
  # generating epsilon
  tausq = 0.01
  epsilon = rnorm(n, mean = 0, sd = sqrt(tausq))
  
  Y = X_1 * beta_1 + X_2 * beta_2+ X_3 * beta_3 + w + epsilon
  
  k = 2
  
  lat_knots = unique(lat)
  lat_knots = lat_knots[seq(1, length(lat_knots), by = k)]
  
  lon_knots = unique(lon)
  lon_knots = lon_knots[seq(1, length(lon_knots), by = k)]
  
  knots = as.matrix(expand.grid(lat_knots, lon_knots))
  
  df = data.frame(coords, Y, X_1, X_2,X_3, w, beta_1, beta_2, beta_3, epsilon)
  
  knots_df = data.frame(knots)
  colnames(knots_df) = c("lat", "lon")
  
  knots_df = merge(knots_df, df, by = c("lat", "lon"), all.x = TRUE, all.y = FALSE)
  
  X = as.matrix(df[, c("X_1", "X_2","X_3")])
  Y_star = knots_df$Y
  X_star = as.matrix(knots_df[, c("X_1", "X_2","X_3")])
  
  m = nrow(knots)
  l = 0
  u = 4
  
  start_time <- Sys.time()
  output = svc::svclm_beta(Y = Y,
                         X = X,
                         s = coords,
                         Y_knots = Y_star,
                         X_knots = X_star,
                         knots = knots,
                         beta_knots_start = matrix(0, nrow = m, ncol = p),
                         phi_beta_start = rep(l + (u - l) / 2, p),
                         sigmasq_beta_start = rep(1, p),
                         tausq_start = 1,
                         phi_beta_proposal_sd = rep(0.5, p),
                         phi_beta_lower = rep(l, p),
                         phi_beta_upper = rep(u, p),
                         mcmc = mcmc)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  gibbs7 = rbind(gibbs7,c(mean(colMeans(output$beta_samples[800:1000, ,1]) -3),mean(colMeans(output$beta_samples[800:1000, ,2]) +5),mean(colMeans(output$beta_samples[800:1000, ,3]) +10), time_taken ))
  X = cbind(1,X)
  start_time <- Sys.time()
 fit <- SVC_mle(y = Y, X = X,locs = coords)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  
  coef7 = rbind(coef7,c(fit$coefficients[2]-3,fit$coefficients[3]+5,fit$coefficients[4]+10,time_taken))
}

gibbs8 = data.frame()
coef8 = data.frame()
### 361 Sample size
for (i in 1:5) {
  lat = seq(0, 9, by = .4)
  lon = seq(0, 9, by = .4)
  
  coords = as.matrix(expand.grid(lat, lon))
  colnames(coords) = c("lat", "lon")
  
  n = nrow(coords)
  p = 3
  
  # generating w
  sigmasq_w = 4
  phi_w = 8
   
  C_w = calc_C_phi(coords, phi_w)
  w = MASS::mvrnorm(1, rep(0, n), sigmasq_w * C_w)
  
  # generating beta's
  sigmasq_1 = 4
  phi_1 = 8
   
  C_1 = calc_C_phi(coords, phi_1)
  beta_1_mean = 3
  beta_1 = MASS::mvrnorm(1, rep(beta_1_mean, n), sigmasq_1 * C_1)
  
  sigmasq_2 = 4
  phi_2 = 8
   
  C_2 = calc_C_phi(coords, phi_2)
  beta_2_mean = -5
  beta_2 = MASS::mvrnorm(1, rep(beta_2_mean, n), sigmasq_2 * C_2)
  
  sigmasq_3 = 4
  phi_3 = 8
   
  C_3 = calc_C_phi(coords, phi_3)
  beta_3_mean = -10
  beta_3 = MASS::mvrnorm(1, rep(beta_3_mean, n), sigmasq_3 * C_3)
  
  # generating X
  X_1 = rnorm(n, 0, 10)
  X_2 = rnorm(n, 10, 100)
  X_3 = rnorm(n, 100, 1000)
  
  # generating epsilon
  tausq = 0.01
  epsilon = rnorm(n, mean = 0, sd = sqrt(tausq))
  
  Y = X_1 * beta_1 + X_2 * beta_2+ X_3 * beta_3 + w + epsilon
  
  k = 2
  
  lat_knots = unique(lat)
  lat_knots = lat_knots[seq(1, length(lat_knots), by = k)]
  
  lon_knots = unique(lon)
  lon_knots = lon_knots[seq(1, length(lon_knots), by = k)]
  
  knots = as.matrix(expand.grid(lat_knots, lon_knots))
  
  df = data.frame(coords, Y, X_1, X_2,X_3, w, beta_1, beta_2, beta_3, epsilon)
  
  knots_df = data.frame(knots)
  colnames(knots_df) = c("lat", "lon")
  
  knots_df = merge(knots_df, df, by = c("lat", "lon"), all.x = TRUE, all.y = FALSE)
  
  X = as.matrix(df[, c("X_1", "X_2","X_3")])
  Y_star = knots_df$Y
  X_star = as.matrix(knots_df[, c("X_1", "X_2","X_3")])
  
  m = nrow(knots)
  l = 0
  u = 4
  
  start_time <- Sys.time()
output = svc::svclm_beta(Y = Y,
                         X = X,
                         s = coords,
                         Y_knots = Y_star,
                         X_knots = X_star,
                         knots = knots,
                         beta_knots_start = matrix(0, nrow = m, ncol = p),
                         phi_beta_start = rep(l + (u - l) / 2, p),
                         sigmasq_beta_start = rep(1, p),
                         tausq_start = 1,
                         phi_beta_proposal_sd = rep(0.5, p),
                         phi_beta_lower = rep(l, p),
                         phi_beta_upper = rep(u, p),
                         mcmc = mcmc)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  gibbs8 = rbind(gibbs8,c(mean(colMeans(output$beta_samples[800:1000, ,1]) -3),mean(colMeans(output$beta_samples[800:1000, ,2]) +5),mean(colMeans(output$beta_samples[800:1000, ,3]) +10), time_taken ))
  X = cbind(1,X)
  start_time <- Sys.time()
 fit <- SVC_mle(y = Y, X = X,locs = coords)
  end_time <- Sys.time()
  time_taken <- as.numeric(difftime(end_time, start_time, units = "secs"))
  
  coef8 = rbind(coef8,c(fit$coefficients[2]-3,fit$coefficients[3]+5,fit$coefficients[4]+10,time_taken))
}

```


```{r}
combined_table <- rbind(colMeans(gibbs1), colMeans(coef1))

rownames(combined_table) <- c("Gibbs Sampler", "varycoef")
kable(combined_table, caption = "Scenario 1: (2 betas and low priors) - 100 sample size",col.names = c(" Beta 1 - Bias","Beta 2 - Bias", "Time to compute on average"))
```
```{r}
combined_table <- rbind(colMeans(gibbs2), colMeans(coef2))

rownames(combined_table) <- c("Gibbs Sampler", "varycoef")
kable(combined_table, caption = "Scenario 2: (2 betas and low priors) - 500 sample size",col.names = c(" Beta 1 - Bias","Beta 2 - Bias", "Time to compute on average"))
```


```{r}
combined_table <- rbind(colMeans(gibbs3), colMeans(coef3))

rownames(combined_table) <- c("Gibbs Sampler", "varycoef")
kable(combined_table, caption = "Scenario 3: (2 betas and high priors) - 100 sample size",col.names = c(" Beta 1 - Bias","Beta 2 - Bias", "Time to compute on average"))
```


```{r}
combined_table <- rbind(colMeans(gibbs4), colMeans(coef4))

rownames(combined_table) <- c("Gibbs Sampler", "varycoef")
kable(combined_table, caption = "Scenario 4: (2 betas and high priors) - 100 sample size",col.names = c(" Beta 1 - Bias","Beta 2 - Bias", "Time to compute on average"))
```

```{r}
combined_table <- rbind(colMeans(gibbs5), colMeans(coef5))

rownames(combined_table) <- c("Gibbs Sampler", "varycoef")
kable(combined_table, caption = "Scenario 5: (3 betas and high priors) - 100 sample size",col.names = c(" Beta 1 - Bias","Beta 2 - Bias","Beta 3 - Bias", "Time to compute on average"))
```


```{r}
combined_table <- rbind(colMeans(gibbs6), colMeans(coef6))

rownames(combined_table) <- c("Gibbs Sampler", "varycoef")
kable(combined_table, caption = "Scenario 6: (3 betas and low priors) - 500 sample size",col.names = c(" Beta 1 - Bias","Beta 2 - Bias","Beta 3 - Bias", "Time to compute on average"))
```

```{r}
combined_table <- rbind(colMeans(gibbs7), colMeans(coef7))

rownames(combined_table) <- c("Gibbs Sampler", "varycoef")
kable(combined_table, caption = "Scenario 7: (3 betas and high priors) - 100 sample size",col.names = c(" Beta 1 - Bias","Beta 2 - Bias","Beta 3 - Bias", "Time to compute on average"))
```

```{r}
combined_table <- rbind(colMeans(gibbs8), colMeans(coef8))

rownames(combined_table) <- c("Gibbs Sampler", "varycoef")
kable(combined_table, caption = "Scenario 8: (3 betas and high priors) - 500 sample size",col.names = c(" Beta 1 - Bias","Beta 2 - Bias","Beta 3 - Bias", "Time to compute on average"))
```






### Interpretations of results
1. Our model (Gibbs sampler) has high bias compared to varycoef
2. Our model (Gibbs sampler) is quick to compute compared to varycoef

### Next steps
Improve the model Gibbs sampler for better performance (bias)
Add another metric (EES/sec) to compare between two methods